Solving Variational Inequalities and Cone Complementarity Problems in Non-Smooth Dynamics using the Alternating Direction Method of Multipliers

### 2021-02-16 2021/ADMM_journal/responses_to_reviewers.tex

\documentclass[final,12pt]{article}

% For journal layout:
% \documentclass[final,1p,times]{elsarticle}
% \documentclass[final,1p,times,twocolumn]{elsarticle}
% \documentclass[final,3p,times]{elsarticle}
% \documentclass[final,3p,times,twocolumn]{elsarticle}
% \documentclass[final,5p,times]{elsarticle}
% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% Use packages ---------------------------------------------
\usepackage[english]{babel}
\usepackage{amssymb} 
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{color}
\usepackage{subcaption}

%% Customization --------------------------------------------
% New commands Tasora 
\def\vect#1{{\boldsymbol{#1}}}
\def\quat#1{{\boldsymbol{#1}}}
\def\conj{\hskip0.1em^*} % the matrix as [R(a)] , with letter between two (autoresizing) square brackets
\def\amatr#1{{#1}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

% New commands Fusai 
\newcommand{\hl}[1]{\textcolor{red}{#1}}

% comandi per risposte a reviewers  etc.

% METTERE LA SEGUENTE LINEA COME COMMENTO, O TOGLIERLA, PER VEDERE LA LETTERA 
% DI RISPOSTA SENZA LE NOSTRE ANNOTAZIONI Giovanni:... Alessandro:.. Dario...  
% E LASCIARE SOLO COMMENTI E RISPOSTE

\def\showcomments{1}

\definecolor{mycommentcolor}{rgb}{0.23, 0.45, 0.6}
\definecolor{alecolor}{rgb}{0.0, 0.6, 0.0}

\def\reviewercomment#1{{\vskip4mm \color{mycommentcolor} \textit{#1} \vskip2mm}}

\def\alessandro#1{
\ifx\showcomments\undefined
\else
    {\color{alecolor} \small \begingroup \leftskip=1.5cm \noindent Alessandro: #1 \par \endgroup}
\fi
}

% Title
\title{ \textbf{Responses to Reviewer's  on:} \\
Solving Variational Inequalities and Cone Complementarity Problems in Non-Smooth Dynamics using the Alternating Direction Method of Multipliers}

\author{A. Tasora, D. Mangoni, S. Benatti, R. Garziera}

%% Document ----------------------------------------------
\begin{document}

\maketitle

% Sections

We would like to thank the reviewers for their thoughtful comments and
efforts towards the improvement of our manuscript. In the following document we include
our replies to their comments, where the light text color (in italic font) is used
for the original comment from the reviewers.

\section{Reviewer 1}

We thank the Reviewer for the useful comments. 
In the following we will address his requests and suggestions.

\reviewercomment{
1. The references cited for the Measure Differential Inclusion (MDI) include a CISM course and a
conference. These are not resources that are easily accessible to everyone. I would strongly
recommend that an accessible resource be cited in their place. Same holds for references 3,
10, 18, 20, 21, 35.
}

We replaced references [1] and [2] with journal articles, as suggested by the Reviewer. We kept reference [3] as is, because it is not a conference but a quite popular book. We kept also reference [10] because the ACM SIGGRAPH conferences are important and their proceedings are always accessible. The article [18] is a journal paper, available online nowadays (some difficulty to access it depends on the fact that Morrocco, the name of the author, is spelled Morroco in some databases). We left other references to conferences when proceedings are accessible and there are no alternative versions on journals.

\reviewercomment{
2. DOI’s are probably not useful for older papers that are already published. I would suggest
uniformity by mentioning DOI for all or none.
}

I suppose unnecessary DOIs will be managed later, by the typesetting department at Wiley, because in the "`Author guidelines"' web page of NME they say \textit{Where possible the DOI for the reference should be included at the end of the reference}.

\reviewercomment{
3. The authors cite Ref [3] as a source to discussion complementarity problems. But this is not a
resource again available easily and I would recommend citing a more resourceful reference.
}

We added more references to VIs and complementarity problems in that paragraph, mentioning two other popular references.

\reviewercomment{
4. I would recommend a nomenclature and abbreviation listing at the start considering that there
are several abbreviations employed in the text.
}

We added a listing of the abbreviations at the beginning of the paper.

\reviewercomment{
5. In page 3, the authors specify that the CCP and NCP approaches are not suitable for
applicability with finite elements. This is a very generic statement. I would like to highlight two
recent works where CCP has been implemented with ANCF elements and in a Finite Element
framework. These works also compare the performance with standard methods (like penalty)
that are often employed. \\
...
}

We thank the Reviewer for pointing out these recent papers. We added them to the list of references and we improved the discussion (see revised interoduction). 
We note, in passing, that the CCP methods outlined in the two references above use a Schur matrix in the form $N=D^TM^{-1}D$, hence without the introduction of tangent stiffness and damping matrices as jacobians of internal forces (something that could be needed when the stiffness of the elements is too large to be handled by the time stepping method without running into instability). This was the rationale of our paragraph: once the Schur matrix becomes too dense to be handled efficiently, one should avoid assembling it at all - and in the current paper we avoid building it, in fact. This is also discussed later, between Formula 31 and Theorem 1.

\reviewercomment{
6. From the introduction and literature review, the aim of the authors is unclear. Is this paper
oriented towards computer graphics or mechanics-related applications? Most of the
discussions pertaining to the available literature is aimed at computer graphics applications.
While non-penetration is an important condition, accurate retrieval of the contact stresses is
also particularly important for mechanics-related applications. This is not discussed in the
literature review nor in the results.
}

The paper is oriented more towards mechanical application, indeed. The fact that literature is mostly targeting computer graphics depends on the fact that complementarity-based approaches are very robust - a quality that triggered lot of interest in the computer graphics community in the last decade. At the other hand, the CG community is often not interested in results that are precise at an engineering level (as they can accept visually-plausible results with low fidelity), so this is stimulating our research on methods that could satisfy the needs of both words: the need of stability and robustness, from the CG field, and the need of precision from the engineering field. Another point that facilitates the problem in the CG area is the fact that CG problems with contacts and deformable parts often involve jelly-like objects with low stiffness (rods, hairs, tissues, flesh), whereas in the engineering area we often deal with steel, aluminium and other structures with high stiffness. We added a discussion at the beginning of the Benchmark section.

\reviewercomment{
7. The “Remark” before and after Definition 6 is the same. Is there a particular reason for the
repetition?
}

Thank you for noticing this. We removed the repetition.

\section{Reviewer 2}

\reviewercomment{
The manuscript focuses on accelerating the solution of the frictional contact problem for rigid and flexible body dynamics. The authors work with an established way of modeling the frictional contact problem, and subsequently carrying out its time discretization. The resulting CCP is solved by means of a new technique, which draws on work done in optimization in relation to the Alternating Direction Method of Multipliers method.
}

We thank the Reviewer for his notes, that are helping us to increase the quality of the paper. Also we thank the Reviewer because of the many stylistic improvements that he listed, about grammar, syntax and style.

 We address his requests and comments in the following paragraphs:

\reviewercomment{
- without adding $C_\epsilon$ in Eq. 33(a), would this approach lead to a unique set of frictional contact forces? Or is there a possibility of finding an infinite number of compatible solutions? Is adding $C_\epsilon$ eliminating this possibility?
}

Correct, the $C_\epsilon$ has an effect toward the regularization of the system, therefore a nonzero $C_\epsilon$ guarantees that there is a unique set of solutions (the mechanical equivalence is that there are both normal and tangential springs at each contact, thus eliminating the redundancy of forces that would arise even in very simple problems like a three leg table over a flat plane if there is static friction). For zero $C_\epsilon$, one cannot rule out the case that $\rho$ becomes close to zero because of the $\rho$ automatic adjustment during the iteration, and in that case one could approach singularity (note that this would also correspond to the pivot-breakdown of the solver if a direct solver is used).

\reviewercomment{
- the CCP approach advocated here has some numerical artifacts, e.g., if a brick moves really fast on ground, there will be a lifting force that reduces the normal force at the brick-ground interface. Is the ADMM approach advocated here good enough to deal with other frictional contact models that lead to optimization problems whose solution eschews the said lift-off artifact?
}

This is an interesting point. The CCP approach we refer to, is based on a convexified model where one can experience the "lifting" effect. \footnote{We note, in passing, that this lift off is balanced by a stabilization coefficient that makes the sliding object to float with a gap of height $h\approxeq v \mu h$, where $v$ is tangential speed, $\mu$ is friction coeff. and $h$ is time step, so it is negligible for small time steps or small speeds.} The proper way to get rid of the lifting effect would be to use the original non-convexified complementarity problem. Using the De Saxce bipotential formulation of the generic non-associated Coulomb friction, CCP of Eq.13 looses the friendly affine structure of the 
$\vect{u}_\epsilon =  N \vect{\gamma}_\epsilon + \vect{r}_\epsilon$ 
term, and one would use rather 
$\vect{u}_\epsilon =   N \vect{\gamma}_\epsilon + \vect{r}_\epsilon + \tilde{\vect{u}}_\epsilon(\vect{\gamma}_\epsilon)$ where $\tilde{\vect{u}}_\epsilon(\vect{\gamma}_\epsilon)$ is a term that depends quadratically on the velocity of tangential sliding. However, this done, the resulting CCP 
$-\Upsilon^{\circ} \ni \vect{u}_\epsilon \bot \vect{\gamma}_\epsilon \in \Upsilon$ would not lead any more to a convex optimization problem. Although the paper [31] (Anderson Acceleration etc., Ouyang et al. 2020) speculates that ADMM could solve non-convex problems, this has been proved only empirically and for limited cases, so we prefer not to discuss the non-convex case in our current manuscript: it would be an interesting extension though, and we thank the Reviewer for the suggestion.

\reviewercomment{
Suggested list of corrections, small improvements:
- Abstract: “multi-flexible-body problems” – perhaps “flexible multi-body problems”?
}

$\rightarrow$ Corrected.

\reviewercomment{
- Abstract: “In sake of computational performance” – “In order to improve computational performance”?
}

$\rightarrow$ Corrected.

\reviewercomment{
- P.1.: “In sake of better numerical performance and stability” – “In search of better numerical performance and stability”. NOTE: the authors should revisit the use of “in sake” since it’s not a proper language construct. It shows up in several places in the manuscript
}

$\rightarrow$ Corrected, also in other places. Thank you for helping us to improve the language of the manuscript.

\reviewercomment{
- P2: “stimulated many researches on” – “stimulated research on” ?
}

$\rightarrow$ Corrected.

\reviewercomment{
- P2: “algorithm, however” – maybe start a new sentence “algorithm. However,…”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P2: “sub iterations” – “sub-iterations”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P2: “is even more critical if odd mass” – “is even more challenged if odd mass”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P2: “happen when simulating” – “happens when simulating”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P2: “in form” – “in the form”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P2: “newton methods” – “Newton methods”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P2: “This requires a simplifying assumption about the Coulomb friction being associative instead of non-associative” – can you please explain what this statement means, perhaps even in one sentence in less, in the manuscript?
}

Thank you for the suggestion. We added two or three lines explaining this point:  
\textit{This artificial associativity of the friction model means that the relative sliding velocity in a contact point will be restricted to the dual cone of the Coulomb friction cone, whereas the original model has a more relaxed condition, assuming that such velocity could be even parallel to the surface: the practical consequence is that grazing motion cause an artificial lift-off - however this artifact can be attenuated thanks to the introduction of a stabilization term [cite{ani04}]}

\reviewercomment{
- P2: “offer a superior convergence respect” – “offer superior convergence with respect”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P2: “in fact they require” – “in that they require”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P2: “and cast the original” – “and casts the original”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P3: “they compares”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P3: “on the top of this method we” – “building off this method, we”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P3: “of such ADMM is that” – “of the approach is that”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P3: “such linear system most often” – “the linear system most often”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P3: “In our embodiment” – “In the approach proposed herein”
}

$\rightarrow$ Corrected.\\

\reviewercomment{
- P4: “Remark. The indicator function of any closed set is lower semicontinuous…” – this shows up twice
}

$\rightarrow$ Corrected.

\reviewercomment{
- P4 (footnote): “the plastic flow might deviate respect to the” – “the plastic flow might deviate with respect to the”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P9: “We experienced that it is possible to use a” – “We experimented with the idea of using a”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P11: “in fact, after ADMM computes” – “In fact, after ADMM computes”
}

$\rightarrow$ Corrected.

\reviewercomment{
- P12: “Chrono open-source library” – provide reference to this library. Also to Eigen and Pardiso
}

Two references to papers about the mentioned libraries have been added.

\vskip15mm

Best regards,

\vskip10mm

the Authors

\end{document}
\endinput
%% End document -----------------------------------------

### 2021-02-17 2021/ADMM_journal/author_v1.tex
old revision for reference

% to typeset URLs, URIs, and DOIs
\usepackage{url}
%\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{cleveref}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{siunitx}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{tabularx}
%\usepackage[ruled,vlined]{algorithm2e}  % ruled,vlined,linesnumbered % NOT WORKIGN IN IJNME, they use algorithmicx
\usepackage{array}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{xcolor}
%\usepackage[title]{appendix}
\usepackage{empheq}
\usepackage{framed} % Framing content
\usepackage{nomencl} % Nomenclature package
\makenomenclature
\setlength{\nomlabelwidth}{2cm}
\setlength{\nomitemsep}{-0.1cm} % Baseline skip between items

@@

\abstract[Summary]{
This work presents a numerical method for the solution of variational inequalities arising in non-smooth flexible multi-body problems that involve set-valued forces. For the special case of hard frictional contacts, the method solves a second order cone complementarity problem. We ground our algorithm on the Alternating Direction Method of Multipliers (ADMM), an efficient and robust optimization method that draws on few computational primitives. In order to improve computational performance, we reformulated the original ADMM scheme in order to exploit the sparsity of constraint jacobians and we added optimizations such as warm starting and adaptive step scaling. The proposed method can be used in scenarios that pose major difficulties to other methods available in literature for complementarity in contact dynamics, namely when using very stiff finite elements and when simulating articulated mechanisms with odd mass ratios. The method can have applications in the fields of robotics, vehicle dynamics, virtual reality, and multi-physics simulation in general.
}

@@

In search of better numerical performance and stability, even in case of large time steps, we rather formulate the original non-smooth dynamical problem as a Measure Differential Inclusion (MDI). MDIs were pioneered by Moreau in the eighties, among others, aiming at a viable approach to the simulation of mechanical systems with hard contacts
\cite{Moreau1983,moreau1988topics}. 
Being differential inclusions, they can directly embed set-valued force laws, such as the Coulomb friction model, but they also generalize to the case where velocity is assumed to be a (possibly) discontinuous function of bounded variation, therefore they admit impulsive events.

MDIs can be solved using special time stepping methods that offer high stability and robustness, however this comes at the cost of solving a Complementarity Problems (CP) per each time step \cite{acary2008numerical}. These CPs correspond to the broader class of Variational Inequalities (VI) \cite{Kinderleher1980,Harker1990}.

In general, the solution of VIs or CPs represents the major numerical bottleneck in the entire time stepping process of the MDI, especially when dealing with many parts or many contacts. For instance, similar scenarios can happen when simulating packaging devices, ground-machine interaction, rock dynamics, masonry structures and granular materials, where the unknowns of the VIs (contact forces) can amount to thousands or millions. This challenge stimulated research on efficient numerical methods in the last three decades.

Among the former strategies for solving this class of problems, we cite the idea of solving a Linear Complementarity Problem (LCP) per each time step
\cite{StTr95}  % LCP solved via Lemke
.
Once cast as a LCP, the VI problem can be solved by direct LCP solvers such as the Lemke algorithm. However, LCP direct solvers often lead to complex sub-iterations that do not scale well with increasing number of unknowns, hence they are not much used nowadays. Another drawback of casting the problem as a LCP is that they introduce a polyhedral approximation, for example the Coulomb friction cone would be approximated as a faceted pyramid, introducing a numerical source of anisotropy.

At the other end of the spectrum are iterative methods that aim at high robustness and efficiency even at the expense of sacrificing accuracy, since this is often acceptable in fields like real time simulation, video games and virtual reality \cite{Bender2014}.
In these fields, the most used approach nowadays consists in fixed point iterations, similarly to Gauss-Seidel, SOR or Jacobi stationary methods, interleaved with simple projections onto the constraint sets at each iteration
\cite{anitescuTasora2008,massSplittingRichard2012,AleMihaiFriction2013} % P-GS iteration
.
The major issue of these methods is that they exhibit slow convergence when the system contains long kinematic chains or long sequences of contacts, such as when including high stacks of objects \cite{TasoraAnitescuCMAME10}. Their convergence is even more challenged if odd mass ratios are encountered: this often happens when simulating industrial robots, earth-moving machines, complex vehicle suspensions, car drivelines. In these cases the convergence might stall soon, hence if the iterations are truncated in order to meet a time budget in real-time applications, the resulting simulation can be affected by low accuracy - usually detectable in the form of objects that interpenetrate at the point of contact and mechanisms that fall apart. Optimizations like warm starting and sub-stepping can alleviate the drawbacks of these methods \cite{Macklin2019a}.

A more advanced class of solvers is represented by non-smooth Newton methods, such as the one presented in 
\cite{Macklin2019} % Non-smooth Newton Method
, that assume a generic Nonlinear Complemetarity Problem (NCP), again a sub case of a VI, and that enforce the NCP complementarity constraints using non-smooth functions like the Fisher-Burmeister function. In these methods, a generalized non-smooth Newton method can be used to find the zero of the functions, but this comes at the cost of solving a linear system per each iteration.

Recent efforts pushed forward the idea of casting the VI of non-smooth dynamics as a convex optimization problem, namely a Quadratic Program (QP) with convex constraints. This requires a simplifying assumption about the Coulomb friction being associative instead of non-associative. This artificial associativity of the friction model means that the relative sliding velocity in a contact point will be restricted to the dual cone of the Coulomb friction cone, whereas the original model has a more relaxed condition, assuming that such velocity could be even parallel to the surface: the practical consequence is that grazing motion cause an artificial lift-off - however this artifact can be attenuated thanks to the introduction of a stabilization term \cite{ani04}. In this setting, the problem is also a convex Cone Complementarity Problem (CCP). The convexity assumption allows the adoption of numerical methods already available for large scale optimization problems. In \cite{heynIJNME2013}, the Barzilai-Borwein spectral projected gradient is used to this end, and in \cite{hammadTOG2015} we used the Nesterov accelerated projected gradient descend to simulate large scale contact problems.
These are first-order optimization methods, in the sense that they do not require the computation of the Hessian matrix: in fact they rely only on  matrix-by-vector multiplications and inner products (similarly to Krylov subspace methods in the field of linear systems), plus a fast projection operator. Although they offer superior convergence with respect to fixed point iterations, they are not easily applicable to problems involving finite elements. In fact, this class of solvers often perform the matrix-by-vector primitive over a Schur complement matrix, a product of constraint jacobians and inverse of the mass matrix. Since the mass matrix is often (block) diagonal, the Schur complement can be quickly factored at the beginning of the iteration, but if finite elements are used, the Schur matrix would also involve to some extent the inverse of element stiffness matrices and damping matrices, a costly and inefficient operation unless special developments are considered \cite{Francu2017}.
Extending complementarity-based approaches in order to encompass finite elements is important because recent research has shown that complementarity-based formulations for large deformation dynamics with contacts often show superior performance when compared to penalty-based approaches \cite{Yu2020,Bozorgmehri2021}.

Another class of solvers that can draw on the optimization-based approach are Interior Point Methods (IPMs). These solvers share some similarities with the previously mentioned non-smooth Newton methods, in that they require the solution of a saddle-point linear system at each iteration. IPMs offer the best theoretical convergence for the problem at hand \cite{Potra2000}. However, their implementation is complex, and despite the encouraging theoretical properties, in practical scenarios they do not scale well for large problems. One of their drawbacks, also, is that at the present state of knowledge there are no efficient ways to warm-start them \cite{Mangoni2018}. This led us to investigate other methods for constrained optimization problems.

The Alternating Direction Method of Multipliers (ADMM) has been proposed in the 1970s 
\cite{Glowinski75} % [ ADMM invention ]
\cite{Gabay1976} % [ ADMM invention ]
as a practical way to solve constrained optimization problems by iterating over the minimization of two (not necessarily differentiable) objectives.
ADMM introduces auxiliary variables and casts the original problem as the minimization of a separable objective subject to a linear constraint between the original variables and the auxiliary variables. With a proper choice of auxiliary variables, the minimization of the separable function is obtained iterating on two distinct minimization steps of lower complexity. The method iterates updating primal variables, auxiliary variables and dual variables similarly to a fixed point iteration: as such, they compare less favourably to IPMs if one is interested in convergence to high precision, however recent developments showed that in practical scenarios they offer superior speed, scalability and robustness
 \cite{Boyd2011}. % [ ADMM general theory and development ].
These positive properties triggered a recent revival of ADMM and similar operator-splitting methods even as alternatives to IPMs, especially in distributed convex optimization with large-scale problems where moderate precision is acceptable
\cite{Cannon2019}. % [COSMO  conic operator splitting method for large convex problems].

@@

Recent developments addressed the possibility of accelerating the ADMM method.
In 
 \cite{Goldfarb2013} % [ ADMM with acceleration, both convex (non-strongly-convex require step skipping), one differentiable]
an acceleration scheme has been proposed for special cases of symmetric ADMMs, and assuming that one of the two objectives is differentiable.
%
In 
 \cite{Goldstein2014} % [ ADMM and AMA with Nesterov. Convergence ok if both problems are strongly convex (non-strongly-convex require restarting), one is quadratic ].
the Nesterov acceleration method has been proposed for problems where at least one of the two objectives is quadratic, and under the assumption that both objectives are strongly convex; if the latter assumption is not guaranteed, restarting strategies are needed. 
%
In 
 \cite{Kadkhodaie2015} %  [ A2DM2 accelerated ADMM, strongly convex ]
a similar acceleration method has been proposed, with $\mathcal{O}\left(\frac{1}{k^2}\right)$ convergence rate, but without the assumption that one of the objectives must be quadratic.
%
Other types of improvements, based on the Anderson fixed point acceleration, have been put forward in 
 \cite{Zhang2019} % [ ADMM + Anderson acceleration, computergraphics ].
 \cite{Ouyang2020} % [ ADMM + Anderson acceleration, using Douglas-Rachford splitting, computergraphics ]
, showing high performance in computer graphics applications.

@@

In
\cite{Stellato2020} % [OSQP: an operator splitting solver for quadratic programs].
a specialized ADMM method for quadratic problems (OSQP) has been presented: building off this method, we develop a solver that can exploit the nature of non-smooth dynamical problems. An attractive property of the approach is that it requires the solution of a linear system per each iteration, but unlike IPMs, the linear system most often remains unchanged during the iterations, so a factorization can be reused multiple times with a great benefit in terms of speed.

In the approach proposed herein, unknowns are primal variables (velocity measures) and dual variables (impulses in contact points and joints), the convex cone constraints stem from the Coulomb friction law, and the system matrix includes terms from the mass matrices and from the tangent stiffness/damping matrices (hence it is sparse, but not necessarily diagonal).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% NOTE FOR THE TYPESETTING DEPARTMENT: IF THE NOMENCLATURE FRAME
%% CONSUMES TOO MUCH SPACE IN THE ARTICLE PAGES, YOU CAN ALSO REMOVE IT. WE ARE
%% PROPOSING IT, BUT WE CAN ALSO DO WITHOUT IT, IF WE ARE GOING OUT OF THE PAGE BUDGET.

    \nomenclature[10]{ADMM}{Alternating Direction Method of Multipliers}
    \nomenclature[10]{FADMM}{Fast Alternating Direction Method of Multipliers}
    \nomenclature[20]{VI}{Variational Inequality}
    \nomenclature[30]{MDI}{Measure Differential Inclusion}
    \nomenclature[40]{CP}{Complementarity Problem}
    \nomenclature[40]{NCP}{Nonlinear Complementarity Problem}
    \nomenclature[50]{LCP}{Linear Complementarity Problem}
    \nomenclature[60]{CCP}{Cone Complementarity Problem}
    \nomenclature[70]{QP}{Quadratic Program}
    %\nomenclature[80]{IPM}{Interior Point Method}

    \begin{table*}[!t]
       \begin{framed}
         \printnomenclature
       \end{framed}
    \end{table*}

@@

\bibliography{../../bibliography/refsMBS,../../bibliography/refsOPT}

>>>>

% to typeset URLs, URIs, and DOIs
\usepackage{url}
%\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{cleveref}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{siunitx}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{tabularx}
%\usepackage[ruled,vlined]{algorithm2e}  % ruled,vlined,linesnumbered % NOT WORKIGN IN IJNME, they use algorithmicx
\usepackage{array}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{xcolor}
%\usepackage[title]{appendix}
\usepackage{empheq}

@@

\abstract[Summary]{
This work presents a numerical method for the solution of variational inequalities arising in non-smooth multi-flexible-body problems that involve set-valued forces. For the special case of hard frictional contacts, the method solves a second order cone complementarity problem. We ground our algorithm on the Alternating Direction Method of Multipliers (ADMM), an efficient and robust optimization method that draws on few computational primitives. In sake of computational performance, we reformulated the original ADMM scheme in order to exploit the sparsity of constraint jacobians and we added optimizations such as warm starting and adaptive step scaling. The proposed method can be used in scenarios that pose major difficulties to other methods available in literature for complementarity in contact dynamics, namely when using very stiff finite elements and when simulating articulated mechanisms with odd mass ratios. The method can have applications in the fields of robotics, vehicle dynamics, virtual reality, and multi-physics simulation in general.
}

@@

In sake of better numerical performance and stability, even in case of large time steps, we rather formulate the original non-smooth dynamical problem as a Measure Differential Inclusion (MDI). MDIs were pioneered by Moreau in the eighties, among others, in sake of a viable approach to the simulation of mechanical systems with hard contacts
\cite{mor88,Jean1992}. 
Being differential inclusions, they can directly embed set-valued force laws, such as the Coulomb friction model, but they also generalize to the case where velocity is assumed to be a (possibly) discontinuous function of bounded variation, therefore they admit impulsive events.

MDIs can be solved using special time stepping methods that offer high stability and robustness, however this comes at the cost of solving a  Variational Inequality (VI) per each time step. In many cases, VIs correspond to Complementarity Problems (CP) 
 \cite{acary2008numerical}.

In general, the solution of VIs or CPs represents the major numerical bottleneck in the entire time stepping process of the MDI, especially when dealing with many parts or many contacts. For instance, similar scenarios can happen when simulating packaging devices, ground-machine interaction, rock dynamics, masonry structures and granular materials, where the unknowns of the VIs (contact forces) can amount to thousands or millions. This challenge stimulated many researches on efficient numerical methods in the last three decades.

Among the former strategies for solving this class of problems, we cite the idea of solving a Linear Complementarity Problem (LCP) per each time step
\cite{StTr95}  % LCP solved via Lemke
.
Once cast as a LCP, the VI problem can be solved by direct LCP solvers such as the Lemke algorithm, however LCP direct solvers often lead to complex sub iterations that do not scale well with increasing number of unknowns, hence they are not much used nowadays. Another drawback of casting the problem as a LCP is that they introduce a polyhedral approximation, for example the Coulomb friction cone would be approximated as a faceted pyramid, introducing a numerical source of anisotropy.

At the other end of the spectrum are iterative methods that aim at high robustness and efficiency even at the expense of sacrificing accuracy, since this is often acceptable in fields like real time simulation, video games and virtual reality \cite{Bender2014}.
In these fields, the most used approach nowadays consists in fixed point iterations, similarly to Gauss-Seidel, SOR or Jacobi stationary methods, interleaved with simple projections onto the constraint sets at each iteration
\cite{anitescuTasora2008,massSplittingRichard2012,AleMihaiFriction2013} % P-GS iteration
.
The major issue of these methods is that they exhibit slow convergence when the system contains long kinematic chains or long sequences of contacts, such as when including high stacks of objects \cite{TasoraAnitescuCMAME10}. Their convergence is even more critical if odd mass ratios are encountered: this often happen when simulating industrial robots, earth-moving machines, complex vehicle suspensions, car drivelines. In these cases the convergence might stall soon, hence if the iterations are truncated in order to meet a time budget in real-time applications, the resulting simulation can be affected by low accuracy - usually detectable in form of objects that interpenetrate at the point of contact and mechanisms that fall apart. Optimizations like warm starting and sub-stepping can alleviate the drawbacks of these methods \cite{Macklin2019a}.

A more advanced class of solvers is represented by non-smooth newton methods, such as the one presented in 
\cite{Macklin2019} % Non-smooth Newton Method
, that assume a generic Nonlinear Complemetarity Problem (NCP), again a sub case of a VI, and that enforce the NCP complementarity constraints using non-smooth functions like the Fisher-Burmeister function. In these methods, a generalized non-smooth Newton method can be used to find the zero of the functions, but this comes at the cost of solving a linear system per each iteration.

Recent efforts pushed forward the idea of casting the VI of non-smooth dynamics as a convex optimization problem, namely a Quadratic Program (QP) with convex constraints. This requires a simplifying assumption about the Coulomb friction being associative instead of non-associative, a simplification whose artifacts can be attenuated thanks to the introduction of a stabilization term \cite{ani04}. In this setting, the problem is also a convex Cone Complementarity Problem (CCP). The convexity assumption allows the adoption of numerical methods already available for large scale optimization problems. In \cite{heynIJNME2013}, the Barzilai-Borwein spectral projected gradient is used to this end, and in \cite{hammadTOG2015} we used the Nesterov accelerated projected gradient descend to simulate large scale contact problems.
These are first-order optimization methods, in the sense that they do not require the computation of the Hessian matrix: in fact they rely only on  matrix-by-vector multiplications and inner products (similarly to Krylov subspace methods in the field of linear systems), plus a fast projection operator. Although they offer a superior convergence respect to fixed point iterations, they are not easily applicable to problems involving finite elements. In fact, this class of solvers often perform the matrix-by-vector primitive over a Schur complement matrix, a product of constraint jacobians and inverse of the mass matrix. Since the mass matrix is often (block) diagonal, the Schur complement can be quickly factored at the beginning of the iteration, but if finite elements are used, the Schur matrix would also involve to some extent the inverse of element stiffness matrices and damping matrices, a costly and inefficient operation unless special developments are considered \cite{Francu2017}.

Another class of solvers that can draw on the optimization-based approach are Interior Point Methods (IPMs). These solvers share some similarities with the previously mentioned non-smooth Newton methods, in fact they require the solution of a saddle-point linear system at each iteration. IPMs offer the best theoretical convergence for the problem at hand \cite{Potra2000}. However, their implementation is complex, and despite the encouraging theoretical properties, in practical scenarios they do not scale well for large problems. One of their drawbacks, also, is that at the present state of knowledge there are no efficient ways to warm-start them \cite{Mangoni2018}. This led us to investigate other methods for constrained optimization problems.

The Alternating Direction Method of Multipliers (ADMM) has been proposed in the 1970s 
\cite{Glowinski75} % [ ADMM invention ]
\cite{Gabay1976} % [ ADMM invention ]
as a practical way to solve constrained optimization problems by iterating over the minimization of two (not necessarily differentiable) objectives.
ADMM introduces auxiliary variables and cast the original problem as the minimization of a separable objective subject to a linear constraint between the original variables and the auxiliary variables. With a proper choice of auxiliary variables, the minimization of the separable function is obtained iterating on two distinct minimization steps of lower complexity. The method iterates updating primal variables, auxiliary variables and dual variables similarly to a fixed point iteration: as such, they compares less favourably to IPMs if one is interested in convergence to high precision, however recent developments showed that in practical scenarios they offer superior speed, scalability and robustness
 \cite{Boyd2011}. % [ ADMM general theory and development ].
These positive properties triggered a recent revival of ADMM and similar operator-splitting methods even as alternatives to IPMs, especially in distributed convex optimization with large-scale problems where moderate precision is acceptable
\cite{Cannon2019}. % [COSMO  conic operator splitting method for large convex problems].

@@

Recent developments addressed the possibility of accelerating the ADMM method.
In 
 \cite{Goldfarb2013} % [ ADMM with acceleration, both convex (non-strongly-convex require step skipping), one differentiable]
an acceleration scheme has been proposed for special cases of symmetric ADMMs, and assuming that one of the two objectives is differentiable.
%
In 
 \cite{Goldstein2014} % [ ADMM and AMA with Nesterov. Convergence ok if both problems are strongly convex (non-strongly-convex require restarting), one is quadratic ].
the Nesterov acceleration method has been proposed for problems where at least one of the two objectives is quadratic, and under the assumption that both objectives are strongly convex; if the latter assumption is not guaranteed, restarting strategies are needed. 
%
In 
 \cite{Kadkhodaie2015} %  [ A2DM2 accelerated ADMM, strongly convex ]
a similar acceleration method has been proposed, with $\mathcal{O}(\frac{1}{k^2})$ convergence rate, but without the assumption that one of the objectives must be quadratic.
%
Other types of improvements, based on the Anderson fixed point acceleration, have been put forward in 
 \cite{Zhang2019} % [ ADMM + Anderson acceleration, computergraphics ].
 \cite{Ouyang2020} % [ ADMM + Anderson acceleration, using Douglas-Rachford splitting, computergraphics ]
, showing high performance in computer graphics applications.

@@

In
\cite{Stellato2020} % [OSQP: an operator splitting solver for quadratic programs].
a specialized ADMM method for quadratic problems (OSQP) has been presented: on the top of this method we develop a solver that can exploit the nature of non-smooth dynamical problems. An attractive property of such ADMM is that it requires the solution of a linear system per each iteration, but unlike IPMs, such linear system most often remains unchanged during the iterations, so a factorization can be reused multiple times with a great benefit in terms of speed.

In our embodiment, unknowns are primal variables (velocity measures) and dual variables (impulses in contact points and joints), the convex cone constraints stem from the Coulomb friction law, and the system matrix includes terms from the mass matrices and from the tangent stiffness/damping matrices (hence it is sparse, but not necessarily diagonal).

@@

@@

%\bibliographystyle{spmpsci}
\bibliography{../../bibliography/refsMBS,../../bibliography/refsOPT}
